---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(tidyverse.quiet = TRUE)
```
# leapfrog

<!-- badges: start -->
[![R-CMD-check](https://github.com/mrc-ide/leapfrog/workflows/R-CMD-check/badge.svg)](https://github.com/mrc-ide/leapfrog/actions)
[![Codecov test coverage](https://codecov.io/gh/mrc-ide/leapfrog/branch/master/graph/badge.svg)](https://codecov.io/gh/mrc-ide/leapfrog?branch=master)
<!-- badges: end -->

Leapfrog is a multistate population projection model for demographic and HIV epidemic estimation.

The name _leapfrog_ is in honor of [Professor](https://blogs.lshtm.ac.uk/alumni/2018/07/16/obituary-professor-basia-zaba/) Basia [Zaba](https://translate.google.co.uk/#view=home&op=translate&sl=pl&tl=en&text=%C5%BBaba).

*Note: the CCMPP model and implementation of Bayesian Population Reconstruction in TMB previously here have moved to a separate repository: https://www.github.com/mrc-ide/ccmpp.tmb*

## Simulation model

The simulation model is implemented in a header-only C++ library located in [`inst/include/leapfrog-raw.h`](inst/include/leapfrog-raw.h). This location allows the C++ code to be imported in other R packages via specifying `LinkingTo: leapfrog` in the `DESCRIPTION` file.

The simulation model is callable in R via a wrapper function `leafrogR()` created with [Rcpp](https://www.rcpp.org).

## Installation

Install the development version from [GitHub](https://github.com/mrc-ide/leapfrog) via devtools:

``` r
# install.packages("devtools")
devtools::install_github("mrc-ide/leapfrog")
```

## Example

The file `pjnz/bwa2021_v6.13.pjnz` contains an example Spectrum file constructed
from default country data for Botswana in Spectrum v2.13 (December 2021).

Prepare model inputs.

```{r example}
library(tidyverse)
library(leapfrog)

pjnz <- system.file("pjnz/bwa2021_v6.13.pjnz", package = "leapfrog")

demp <- prepare_leapfrog_demp(pjnz)
hivp <- prepare_leapfrog_projp(pjnz)
```

Simulate 'full' age group (single-year age) and 'coarse' age group (collapsed
age groups) models.

```{r simulate_leapfrog}
lsimF <- leapfrogR(demp, hivp, hiv_strat = "full")
lsimC <- leapfrogR(demp, hivp, hiv_strat = "coarse")
```

Compare the HIV prevalence age 15-49 years and AIDS deaths 50+ years. Deaths 50+ 
years are to show some noticeable divergence between the `"full"` and `"coarse"` 
age group simulations.

```{r sim_prev}

prevF <- colSums(lsimF$hivpop1[16:50,,],,2) / colSums(lsimF$totpop1[16:50,,],,2)
prevC <- colSums(lsimC$hivpop1[16:50,,],,2) / colSums(lsimC$totpop1[16:50,,],,2)

deathsF <- colSums(lsimF$hivdeaths[51:81,,],,2)
deathsC <- colSums(lsimC$hivdeaths[51:81,,],,2)

plot(1970:2030, prevF, type = "l", main = "Prevalence 15-49")
lines(1970:2030, prevC, col = 2)

plot(1970:2030, deathsF, type = "l", main = "AIDS Deaths 50+ years")
lines(1970:2030, deathsC, col = 2)
``` 

### Benchmarking 

_Note: The function `devtools::load_all()` invokes `pkgbuild::compile_dll()`,
which executes with default arguemnt `debug = TRUE`. This compiles the package 
with `-O0` compiler optimisation flags. For benchmarking, make sure to install 
the package or compile the DLL with `pkgbuild::compile_dll(debug = FALSE)`._

```{r benchmarking}
library(bench)
library(eppasm)

fp <- eppasm::prepare_directincid(pjnz)

bench::mark(leapfrogR(demp, hivp, hiv_strat = "full"),
            leapfrogR(demp, hivp, hiv_strat = "coarse"),
            eppasm::simmod(fp),
            check = FALSE)
```

## Code design

### Simulation model

The simulation model is implemented as templated C++ code in `inst/include/leapfrog-raw.h`.
This is  the simulation model may be developed as a standalone C++ library
that can be called by other software without requiring R-specific code features.
The code uses header-only open source libraries to maximize portability.



### R functions

The file `src/leapfrogR.cpp` contains R wrapper functions for the model simulation 
via [Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) and 
[RcppEigen](http://dirk.eddelbuettel.com/code/rcpp.eigen.html).


## Development notes

### Simulation model

* The model was implemented using _Eigen::Tensor_ containers. These were preferred
  for several reasons:
  
  * Benchmarking found they were slighlty more efficient than _boost::multi_array_.
  * Column-major indexing in the same order as R
  * Other statistical packages (e.g. TMB, Stan) rely heavily on _Eigen_ so using
    _Eigen_ containers slims the dependencies.

