---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

options(tidyverse.quiet = TRUE)
```
# leapfrog

<!-- badges: start -->
[![R-CMD-check](https://github.com/mrc-ide/leapfrog/workflows/R-CMD-check/badge.svg)](https://github.com/mrc-ide/leapfrog/actions)
[![Codecov test coverage](https://codecov.io/gh/mrc-ide/leapfrog/branch/demproj-only/graph/badge.svg)](https://codecov.io/gh/mrc-ide/leapfrog?branch=demproj-only)
<!-- badges: end -->

Leapfrog is a multistate population projection model for demographic and HIV epidemic estimation.

The name _leapfrog_ is in honor of [Professor](https://blogs.lshtm.ac.uk/alumni/2018/07/16/obituary-professor-basia-zaba/) Basia [Zaba](https://translate.google.co.uk/#view=home&op=translate&sl=pl&tl=en&text=%C5%BBaba).

*Note: the CCMPP model and implementation of Bayesian Population Reconstruction in TMB previously here have moved to a separate repository: https://www.github.com/mrc-ide/ccmpp.tmb*


## Branch _demproj-only_

The branch _demproj-only_ contains code for only the demographic projection model as a simpler example for developing interfaces with fewer input and output arrays.

## Simulation model

The simulation model is implemented in a header-only C++ library located in [`inst/include/leapfrog-raw.h`](inst/include/leapfrog-raw.h). This location allows the C++ code to be imported in other R packages via specifying `LinkingTo: leapfrog` in the `DESCRIPTION` file.

The simulation model is callable in R via a wrapper function `leafrogR()` created with [Rcpp](https://www.rcpp.org).

## Installation

Install the development version from [GitHub](https://github.com/mrc-ide/leapfrog/tree/demproj-only) via devtools:

``` r
# install.packages("devtools")
devtools::install_github("mrc-ide/leapfrog@demproj-only")
```

## Example

The file `bwa_demproj-only_spectrum-v6.13_2022-02-12.PJNZ` contains an example Spectrum file constructed
from default country data for Botswana in Spectrum v2.13 (December 2021) with _only DemProj_ activated.

Prepare model inputs.

```{r example}
library(leapfrog)
pkgbuild::compile_dll(".", debug = FALSE)
devtools::load_all(".")

pjnz <- "tests/testdata/spectrum/v6.13/bwa_demproj-only_spectrum-v6.13_2022-02-12.PJNZ"

demp <- prepare_leapfrog_demp(pjnz)
```

Simulate 'full' age group (single-year age) and 'coarse' age group (collapsed
age groups) models.

```{r simulate_leapfrog}
lsim <- leapfrogR(demp)
```

Compare population, deaths, and births to Spectrum outputs. The _leapfrog_ model does not produce birth or death outputs for the baseline year.

```{r demproj_outputs}
specres <- eppasm::read_hivproj_output(pjnz)

plot(1970:2030, colSums(lsim$totpop1,,2), type = "l", main = "Total population")
lines(1970:2030, colSums(specres$totpop,,2), col = 2)

plot(1970:2030, colSums(lsim$natdeaths,,2), type = "l", main = "Total non-AIDS deaths")
lines(1970:2030, colSums(specres$natdeaths,,2), col = 2)

plot(1970:2030, lsim$births, type = "l", main = "Births")
lines(1970:2030, specres$births, col = 2)
``` 

## Code design

### Simulation model

The simulation model is implemented as templated C++ code in `inst/include/leapfrog-raw.h`.
This is  the simulation model may be developed as a standalone C++ library
that can be called by other software without requiring R-specific code features.
The code uses header-only open source libraries to maximize portability.



### R functions

The file `src/leapfrogR.cpp` contains R wrapper functions for the model simulation 
via [Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) and 
[RcppEigen](http://dirk.eddelbuettel.com/code/rcpp.eigen.html).


## Development notes

### Simulation model

* The model was implemented using _Eigen::Tensor_ containers. These were preferred
  for several reasons:
  
  * Benchmarking found they were slighlty more efficient than _boost::multi_array_.
  * Column-major indexing in the same order as R
  * Other statistical packages (e.g. TMB, Stan) rely heavily on _Eigen_ so using
    _Eigen_ containers slims the dependencies.

